### 2025-09-29 daj-revision2/referee-response.txt
more referee edits

I am not fully happy with the revision. Most of my comments and sugges
tions have been taken into account, but not all. Several have been ignored for a
good reason (14 15 20 were completely wrong and 7 was perhaps not relevant),
but it seems that the authors have not got what I meant in my comments 34
and 35.

In 34, I was just asking that you explain in the text of your paper how
Corollary 5.12 improves Theorem 7.14 in [JNV+20a]. What you wrote The
statement is stronger than Theorem 7.14 in [JNV+20a], which involves a game
that has a question with long answers is exactly the kind of thing I was asking.
Would you agree to add it to the text?

A: Ok, we added a sentence right before the statement of Corollary 5.12

In 35, I was just suggesting a different way of phrasing your proof. I guess
that I was implicitely complaining that you were reproducing again the same
argument that you had written many times earlier (and that I called “Lemma 1”
in my first report). And that it would be kind to the reader if you encapsulated
the idea as a lemma so that you can only refer to it each time you use it. Now
that you did it in you new Lemma 2.7, you’d better use it rather than reproving
it. See 21 below.

A: [Michael, you had looked at this. Can you check the referee comment 21 and the lemma they propose (see the pdf file for that) and decide if we want to reformulate as they suggest?]

More importantly, the comments 9 10 13 14 19 21 have not been fully ad-
dressed, or resulted in adding new inaccuracies (see points 2–5, 8, 7, 10, 18). In
particular I fully disagree with your reply to 21, see point 18 below.
Here are some comments. Some are new and some are about previous com-
ments.

1. page 9, Definition 2.4. You made the choice to only discuss what is usually
called flexible stability and not the more restrictive notion where you do
not change de dimension/algebra of the representation. This choice is
understandable because flexible stability is much more common that non-
flexible stability, but maybe it’s worth pointing out the stability results
that are true for the stronger form of non-flexible stability. In particular,
this seems to be the case for Theorem 4.1. This deserves to be written.

A: [Michael, are you able to do this? I can do it but I will have to dig out the results one by one as I am not so familiar]

2. page 9, Claim 2.3. There are squares missing. (3) is certainly correct
(perhaps with a different constant than 3), but I did not follow, in the
proof, how you take into account the renormalization of the trace on N .
(4) is obvious with 5 replaced by 1, because by the trace property, if
X = Ui − w∗Viw,
∥X∥τ = ∥X∗∥τ = ∥U −1
i − w∗V −1
i w∥.

A: Thanks, you are correct. Actually the claim was not needed so we removed it.

3. page 10, Lemma 2.5: δ sould be √δ. The fact that you do not put paren-
theses is misleading, it could be understood as (Pr
i=1 ∥...∥τ ) + 2δ. Put
parentheses in the sum, or write (r + 1)√δ + P|r|
i=1 ∥...∥τ .

A: Corrected

4. page 12, (7): μS is not a probability measure.

A: This was a typo; we had meant to include the scaling with |r| as in your previous report.

5. page 12, proof of Lemma 2.7. You are not proving the lemma: your
definition of almost-homomorphism was in terms of E∥ϕ(r) − 1∥2
τ , whereas
you prove something about E∥ϕ(r) − 1∥τ .

A: Corrected

6. page 20, line 6: the phrasing “to guarantee that Ri is independent of the
order...” is misleading. Probably what you mean it that the presentation
(20) defines the same group, whatever the order in which the xi are mul-
tiplied. But as a relation, the order always matters, even after adding the
relations R′
ijj′ . For example, the notion of (ε, μ)-almost homomorphism
changes is the order changes.

A: Right, we corrected this

7. This is really unimportant, but I meant that the two groups in the state-
ment of the lemma are only isomorphic, and I kind of agreed that the two
groups in the proofs were equal.

A: [T: He is referring to Lemma 3.1. I don't really get it. We should say that Z_2^k and the presentation are isomorphic, not equal? Then we need to update this all over. Michael?]

8. page 20, Definition 3.3 and above. You write “we state the definition for
the case of a general prime power q” but then there is no q, just 2. So
both occurancies of 2 should be replaced by q.

A: Right, corrected

9. page 22, line 1-2: you should write down the assumption you have on d
(for d ≥ q, the dimension you write is clearly incorrect).

A: added

10. page 24, line 8. You could say more clearly that you fix ti one and for all,
and that the way they are chosen has no impact on the statements (it was
not clear to me whether they are allowed to change or not, and how they
are chosen).

A: Yes, added

11. page 24, line 3 after Figure 3: is the formula for M correct? I count
qm ×m×q ×q +qm ×2t ×2t = (m+1)qm+2, so either I did not understand
what the tester implied by the figure means, or there is a typo.

A: You are right, this hadn't been updated from an earlier change. We fixed it.

12. page 24, line -7: k = t(d + 1)m?

A: [T: This is what we write, what is the question?]

13. page 25, Theorem 4.1. There is a typo in the formula for δ(ε) (your
formula gives δ(ε) ≥ C; this form of the theorem is of course true, but not
very impressive).

A: Yes, the brackets had been closed in the wrong way in the latex; thanks.

14. page 25, line 1 in the proof of Theorem 4.1: μS = μR.

A: Corrected

15. page 25, Claim 4.2: do you need Claim 4.2? Corollary 2.15 that you use
later does not require that ϕ takes its values in the hermitian unitaries.

A: You're right, these relations are supserseded by the "Hadamard" relations and were not necessary. We removed them from the presentation, removed Claim 4.2, and adapted the rest of the text.

16. page 26, Claim 4.4: why the √ε? Claim 4.3 exactly says that ϕ is
(δ, μS )-close to U with δ = O(ε), and Lemma 2.7 (or its proof) says that
E∥[Uu,α, Uv,α′ ] − [ϕ(su,α), ϕ(sv,α′ )]∥2 ≤ 16δ for any probability measure
on pairs whose marginals are μS .

Yes this was a typo in the last line of the proof, we corrected it.

17. page 28, line 3: what you write is not true for a fixed u, but it is true on
average over u.

A: Corrected

18. page 29, line -5– -1: the conventions for the notation poly here (given
in footnote 7) do not match those of [JNV+22]. For example, on the
last line of page 23 of [JNV+22], you have an expression of the form
poly(ε, n−1) + e−Ω(r/m2) that becomes, in the conclusion of [JNV+22,
Theorem 4.1] poly(ε, n−1, e−Ω(r/m2)). And similarly all over [JNV+22].
So, the way you prove Theorem 4.1 has two problems: (a) you do not
prove that the presentation given in (22) is stable in the sense of Definition
1.2 (because you do not prove limε δ(ε) = 0) and (2) your convention of
poly contradicts your use of it.
When you write Yes, we are sure that the modulus goes to 0 with epsilon,
I agree that this is true, but where is it proven? I did not find the proof
written, neither in [JNV+22] nor in your submission. I thought that what
is proven in [JNV+22] is with δ(ε) ≤ C(mC + dC )(ε1/C + q−1/C ) for some
universal C, which is C(mC + dC )q−1/C > 0 for ε = 0.
If you want to stick to the convention that limε δ(ε) = 0, one cheap way is
to exploit Lemma 2.6 and [GH17] to explain that the modulus of stabil-
ity in Theorem 4.1 is something like min(C1ε, C2poly(ε, 1/q)) for a good
(=polynomial in the parameters) constant C2 and a bad (=exponential in
the parameters) constant C1.
19. page 29, line -2: in [JNV+22], there is an assumption r ≥ 12mt (which
in the notation here becomes r ≥ 12m(d + 1)). Do you have assumptions
that ensure that Ω(tm2) satisfies this inequality?
20. page 29, line -1: you could explain why the existence of Gc implies the
conclusion of the theorem: write the homomorphism that is δ(ε)-close to
ϕ.
21. Here is what I had in mind in my comment 35.
Lemma 1. Assume that G = ⟨S, R⟩ is (δ, μR, μS )-stable, and let νR be
another probability measure on R such that νS = μS .1 Then every (ε, μR)-
almost representation is a (CEr∼ν |r|2δ(ε), νR)-almost representation.2
Proof. By the definition of stability, an (ε, μR)-representation is (δ(ε), μS )-
close to a representation, that is (δ(ε), νS )-close to a representation (νS =
νR). Lemma 2.7 implies that it is a (Cδ(ε), νR)-almost representation.
Let us deduce the new inequality in Corollary 2.8:
∀ε ≤ 1
k − 1 , δk(ε) ≥ kε
C′ .
Let G = ([k], E) be a perfect matching on k vertices, consider νR the
uniform probability measure on {[xi, xi] | ij ∈ E}. Then νS = μS is the
uniform probability measure {x1, . . . , xk}.
1We use the notation from your remark 2.8.
2C is the universal constant in your Lemma 2.7.
3
Consider A the map from Claim 4.2. It is a ( 1
k−1 , μR)-almost represen-
tation. So Φ : xi 7 → (Ai, 1) is a (ε, μR)-almost representation if M ⊕ C
is equipped with the trace τ ε(x, λ) = (k − 1)ετ M(x) + (1 − (k − 1)ε)λ.
Moreover, for every edge ij, ∥[Φ(xi), Φ(xj )] − 1∥2
τ ε = 2(k − 1)ε, and the
lemma implies
2(k − 1)ε = 1
|E|
X
ij∈E
∥[Ai, Aj ] − 1∥2
τ ε ≤ C16δk(ε)